Demes: a standard format for demographic models

### 2022-08-10 paper.tex
Remove line break in abstract

\begin{abstract}
Understanding the demographic history of populations is a
key goal in population genetics, and with improving methods
and data, ever more complex models are being proposed and tested.
Demographic models of current interest
typically consist of a set of discrete populations,
their sizes and growth rates, and continuous and pulse migrations
between those populations over a number of epochs, which can require
dozens of parameters to fully describe. There is currently
no standard format to define such models, significantly
hampering  progress in the field. In particular, the important
task of translating the model descriptions in published
work into input suitable for population genetic simulators is labor intensive
and error prone.
We propose the Demes data model and file format,
built on widely used technologies,
to alleviate these issues. Demes provides a well-defined and unambiguous
model of populations and their properties that is straightforward to
implement in software, and a text file format that is designed for
simplicity and clarity.
We provide thoroughly tested implementations of Demes parsers
in multiple languages including Python and C,
and showcase initial support in several simulators and inference
methods.
An introduction to the file format and a detailed specification are available at:\\
\url{https://popsim-consortium.github.io/demes-spec-docs/}.
\end{abstract}

@@

\begin{figure}
    \centering
    \includegraphics{fig/showcase}
    \caption{
        \textbf{Illustration and simulation using Demes.}
        (A)
        Using an inferred demographic model from \citet{tennessen2012evolution}
        specified as a YAML file in Demes format (Figure~\ref{fig:tennessen}), we
        used \demesdraw\ to visualize the demographic model (note the recent
            exponential growth resulting in present-day population sizes that
        greatly exceed those in the past).
        We then used \msprime\ to simulate genomic data for 20 genome copies
        sampled from the two contemporary populations, and we used \moments\
        to compute the expected joint site-frequency spectrum for the same
        sample sizes (Figure~\ref{fig:tennessen-simulation}).
        (B, C) We compared the single-population SFS in each population, showing
        agreement between the simulation methods.
        (D) Python code snippets of the interactions between \demes\ and the simulation
        software. An extended script to compute the SFS shown in (B) and (C) is
        given in Figure~\ref{fig:tennessen-simulation}.
    }
    \label{fig:showcase}
\end{figure}

@@

\begin{figure}[h!]
    \begin{tcolorbox}
        % NOTE: line highlights may need adjusting after changing this code
        \inputminted[fontsize=\scriptsize,numbersep=5pt,highlightlines={9,19,52}]{python}{models/tennessen-simulation.py}
    \end{tcolorbox}
    \caption{
        \textbf{Simulation of SFS for the Tennessen model.}
        We first load the demographic model using \demes\ (as \texttt{graph}),
        which can then be used by \msprime\ to create the demographic model used in
        \texttt{msprime.sim\_ancestry()}. The same loaded graph can also be
        passed to \moments\ to compute the expected joint SFS.
        To compare the SFS in Figure~\ref{fig:showcase}, we marginalize the
        joint SFS to obtain the single-population SFS for both AFR and EUR populations.
        Lines interfacing \demes\ and other software are highlighted.
    }
    \label{fig:tennessen-simulation}
\end{figure}

>>>>

\begin{abstract}
Understanding the demographic history of populations is a
key goal in population genetics, and with improving methods
and data, ever more complex models are being proposed and tested.
Demographic models of current interest
typically consist of a set of discrete populations,
their sizes and growth rates, and continuous and pulse migrations
between those populations over a number of epochs, which can require
dozens of parameters to fully describe. There is currently
no standard format to define such models, significantly
hampering  progress in the field. In particular, the important
task of translating the model descriptions in published
work into input suitable for population genetic simulators is labor intensive
and error prone.
We propose the Demes data model and file format,
built on widely used technologies,
to alleviate these issues. Demes provides a well-defined and unambiguous
model of populations and their properties that is straightforward to
implement in software, and a text file format that is designed for
simplicity and clarity.
We provide thoroughly tested implementations of Demes parsers
in multiple languages including Python and C,
and showcase initial support in several simulators and inference
methods.
An introduction to the file format and a detailed specification are available
at \url{https://popsim-consortium.github.io/demes-spec-docs/}.
\end{abstract}

@@

\begin{figure*}
    \centering
    \includegraphics{fig/showcase}
    \caption{
        \textbf{Illustration and simulation using Demes.}
        (A)
        Using an inferred demographic model from \citet{tennessen2012evolution}
        specified as a YAML file in Demes format (Figure~\ref{fig:tennessen}), we
        used \demesdraw\ to visualize the demographic model (note the recent
            exponential growth resulting in present-day population sizes that
        greatly exceed those in the past).
        We then used \msprime\ to simulate genomic data for 20 genome copies
        sampled from the two contemporary populations, and we used \moments\
        to compute the expected joint site-frequency spectrum for the same
        sample sizes (Figure~\ref{fig:tennessen-simulation}).
        (B, C) We compared the single-population SFS in each population, showing
        agreement between the simulation methods.
        (D) Python code snippets of the interactions between \demes\ and the simulation
        software. An extended script to compute the SFS shown in (B) and (C) is
        given in Figure~\ref{fig:tennessen-simulation}.
    }
    \label{fig:showcase}
\end{figure*}

@@

\begin{figure}[h!]
    \begin{tcolorbox}
        % NOTE: line highlights may need adjusting after changing this code
        \inputminted[fontsize=\scriptsize,numbersep=5pt,highlightlines={6,14,40}]{python}{models/tennessen-simulation.py}
    \end{tcolorbox}
    \caption{
        \textbf{Simulation of SFS for the Tennessen model.}
        We first load the demographic model using \demes\ (as \texttt{graph}),
        which can then be used by \msprime\ to create the demographic model used in
        \texttt{msprime.sim\_ancestry()}. The same loaded graph can also be
        passed to \moments\ to compute the expected joint SFS.
        To compare the SFS in Figure~\ref{fig:showcase}, we marginalize the
        joint SFS to obtain the single-population SFS for both AFR and EUR populations.
        Lines interfacing \demes\ and other software are highlighted.
    }
    \label{fig:tennessen-simulation}
\end{figure}
